Intelligent Reflecting Surface-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design

### 2021-03-24 main.tex
:pencil: Update notations

			Note that problem~\eqref{op:irs} is not a Semidefinite Programming (SDP) such that relaxing the rank constraint \eqref{co:irs_rank} has no analytical performance guarantee. In Section~\ref{se:performance_evaluation}, we prove through numerical simulations that the solution $\boldsymbol{\Phi}^{\star}$ to \ref{ob:irs} -- \ref{co:irs_sd} is rank-\num{1} under different configurations for all tested channel realizations. It suggests that the rank constraint \eqref{co:irs_rank} can be dropped with negligible loss and $\bar{\boldsymbol{\phi}}^\star$ can be directly obtained through Eigenvalue Decomposition (EVD). In case $\boldsymbol{\Phi}^{\star}$ is not rank-\num{1}, a suboptimal solution can be extracted via the Gaussian randomization method \cite{Huang2010}. Specifically, we decompose $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Sigma}\boldsymbol{U}^H$, generate CSCG random vector $q \in \{1, \dots, Q\}$ by $\boldsymbol{r}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, construct the candidate $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Sigma}^{1/2}\boldsymbol{r}_q\right)}$, and choose the $\bar{\boldsymbol{\phi}}^\star$ that maximizes \eqref{ob:irs} as the solution. Finally, the phase shift of element $l$ is retrieved by $\theta_l^{\star}=\arg(\bar{\phi}_l^\star/\bar{\phi}_{L+1}^\star)$. Problem~\eqref{op:irs} after rank relaxation can be solved by existing optimization tools such as CVX \cite{Grant2013}. The SCA algorithm of passive beamforming is summarized in Algorithm~\ref{al:sca}.
			\begin{algorithm}[!t]
				\caption{SCA: IRS Phase Shift.}
				\label{al:sca}
				\begin{algorithmic}[1]
					\State \textbf{Input} $\beta_2,\beta_4,\boldsymbol{h}_{\mathrm{D},n},\boldsymbol{V}_{n},\boldsymbol{w}_{\mathrm{I/P},n},\rho,\sigma_n,\bar{R},Q,\epsilon$, $\forall n$
					\State Construct $\boldsymbol{V},\boldsymbol{M},\boldsymbol{M}_n,\boldsymbol{C}_{n},\boldsymbol{C}_{\mathrm{I/P},k}$, $\forall n,k$
					\State \textbf{Initialize} $i \gets 0$, $\boldsymbol{\Phi}^{(0)}$
					\State Get $t_{\mathrm{I/P},k}^{(0)}$, $\forall k$ by \eqref{eq:t_k}
					\Repeat
						\State $i \gets i + 1$
						\State Get $\boldsymbol{\Phi}^{(i)}$ by solving \ref{ob:irs} -- \ref{co:irs_sd}
						\State Update $t_{\mathrm{I/P},k}^{(i)}$, $\forall k$ by \eqref{eq:t_k}
						\State Compute $z^{(i)}$ by \eqref{eq:z_irs}
					\Until $\lvert z^{(i)}-z^{(i-1)} \rvert \le \epsilon$
					\State Set $\boldsymbol{\Phi}^{\star}=\boldsymbol{\Phi}^{(i)}$
					\If{$\mathrm{rank}(\boldsymbol{\Phi}^{\star})=1$}
						\State Get $\bar{\boldsymbol{\phi}}^\star$ by EVD, $\boldsymbol{\Phi}^{\star}=\bar{\boldsymbol{\phi}}^\star(\bar{\boldsymbol{\phi}}^\star)^H$
					\Else
						\State Get $\boldsymbol{U},\boldsymbol{\Sigma}$ by EVD, $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Sigma}\boldsymbol{U}^H$
						\State Generate $\boldsymbol{r}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, $\forall q$
						\State Construct $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Sigma}^{1/2}\boldsymbol{r}_q\right)},\boldsymbol{\Phi}_q=\bar{\boldsymbol{\phi}}_q\bar{\boldsymbol{\phi}}_q^H$, $\forall q$
						\State Set $q^{\star}=\arg\max_q{z(\boldsymbol{\Phi}_q)},\bar{\boldsymbol{\phi}}^\star=\bar{\boldsymbol{\phi}}_{q^{\star}}$
					\EndIf
					\State Set $\theta_l^\star=\arg(\bar{\phi}_l^\star/\bar{\phi}_{L+1}^\star)$, $\forall l$, $\boldsymbol{\phi}^{\star}=[e^{j\theta_1^\star},\dots,e^{j\theta_L^\star}]^H$
					\State \textbf{Output} $\boldsymbol{\phi}^{\star}$
				\end{algorithmic}
			\end{algorithm}

@@

			\begin{algorithm}[!t]
				\caption{Modified: IRS Phase Shift.}
				\label{al:lc}
				\begin{algorithmic}[1]
					\State \textbf{Input} $\beta_2,\beta_4,\boldsymbol{h}_{\mathrm{D},n},\boldsymbol{V}_{n},\boldsymbol{w}_{\mathrm{I/P},n},\rho,\sigma_n,Q,\epsilon$, $\forall n$
					\State Construct $\boldsymbol{V},\boldsymbol{M},\boldsymbol{M}_n,\boldsymbol{C}_{n},\boldsymbol{C}_{\mathrm{I/P},k}$, $\forall n,k$
					\State \textbf{Initialize} $i \gets 0$, $\boldsymbol{\Phi}^{(0)}$
					\If{$\rho=0$}
						\State Get $\boldsymbol{\Phi}^{\star}$ by maximizing \eqref{eq:R_irs} w.r.t. \eqref{co:irs_modulus}, \eqref{co:irs_sd}
					\Else
						\Repeat
							\State $i \gets i + 1$
								\State Set $\bar{R}=0$
								\State Get $\boldsymbol{\Phi}^{(i)}$ by solving \ref{ob:irs} -- \ref{co:irs_sd}
								\State Update $t_{\mathrm{I/P},k}^{(i)}$, $\forall k$ by \eqref{eq:t_k}
								\State Compute $z^{(i)}$ by \eqref{eq:z_irs}
						\Until $\lvert z^{(i)}-z^{(i-1)} \rvert \le \epsilon$
						\State Set $\boldsymbol{\Phi}^{\star}=\boldsymbol{\Phi}^{(i)}$
					\EndIf
					\If{$\mathrm{rank}(\boldsymbol{\Phi}^{\star})=1$}
						\State Get $\bar{\boldsymbol{\phi}}^\star$ by EVD, $\boldsymbol{\Phi}^{\star}=\bar{\boldsymbol{\phi}}^\star(\bar{\boldsymbol{\phi}}^\star)^H$
					\Else
						\State Get $\boldsymbol{U},\boldsymbol{\Sigma}$ by EVD, $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Sigma}\boldsymbol{U}^H$
						\State Generate $\boldsymbol{r}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, $\forall q$
						\State Construct $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Sigma}^{1/2}\boldsymbol{r}_q\right)},\boldsymbol{\Phi}_q=\bar{\boldsymbol{\phi}}_q\bar{\boldsymbol{\phi}}_q^H$, $\forall q$
						\State Set $q^{\star}=\arg\max_q{R(\boldsymbol{\Phi}_q)}$ or $q^{\star}=\arg\max_q{z(\boldsymbol{\Phi}_q)}$, $\bar{\boldsymbol{\phi}}^\star=\bar{\boldsymbol{\phi}}_{q^{\star}}$
					\EndIf
					\State Set $\theta_l^\star=\arg(\bar{\phi}_l^\star/\bar{\phi}_{L+1}^\star)$, $\forall l$, $\boldsymbol{\phi}^{\star}=[e^{j\theta_1^\star},\dots,e^{j\theta_L^\star}]^H$
					\State \textbf{Output} $\boldsymbol{\phi}^{\star}$
				\end{algorithmic}
			\end{algorithm}

			Similar to Algorithm~\ref{al:bcd}, we iteratively update the passive beamforming $\boldsymbol{\phi}$ by Algorithm~\ref{al:lc}, the active beamforming $\boldsymbol{b}_{\mathrm{I/P},n}$, $\forall n$ by equation \eqref{eq:b}, and the waveform amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ by \eqref{eq:wf}, \eqref{eq:smf} for TS or \eqref{eq:s_i}, \eqref{eq:s_p} for PS, until convergence.
		\end{subsection}
	\end{section}

@@

		We first evaluate the performance of Algorithm~\ref{al:sca} under SDR. It is demonstrated that $\boldsymbol{\Phi}^{\star}$ is rank-\num{1} for all tested channel realizations with different $M$, $N$ and $L$, such that $\boldsymbol{\phi}^{\star}$ can be directly obtained through EVD and the Gaussian randomization step is not required. Therefore, we claim Algorithm~\ref{al:sca} converges to stationary points of problem~\eqref{op:irs} without performance loss.

>>>>

			Note that problem~\eqref{op:irs} is not a Semidefinite Programming (SDP) such that relaxing the rank constraint \eqref{co:irs_rank} has no analytical performance guarantee. In Section~\ref{se:performance_evaluation}, we prove through numerical simulations that the solution $\boldsymbol{\Phi}^{\star}$ to \ref{ob:irs} -- \ref{co:irs_sd} is rank-\num{1} under different configurations for all tested channel realizations. It suggests that the rank constraint \eqref{co:irs_rank} can be dropped with negligible loss and $\bar{\boldsymbol{\phi}}^\star$ can be directly obtained through eigen decomposition. In case $\boldsymbol{\Phi}^{\star}$ is not rank-\num{1}, a suboptimal solution can be extracted via the Gaussian randomization method \cite{Luo2010}. Specifically, we decompose $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Lambda}\boldsymbol{U}^H$, generate CSCG random vector $q \in \{1, \dots, Q\}$ by $\boldsymbol{\xi}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, construct the candidate $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Lambda}^{1/2}\boldsymbol{\xi}_q\right)}$, and choose the $\bar{\boldsymbol{\phi}}^\star$ that maximizes \eqref{ob:irs} as the solution. Finally, the phase shift of element $l$ is retrieved by $\theta_l^{\star}=\arg(\bar{\phi}_l^\star/\bar{\phi}_{L+1}^\star)$. Problem~\eqref{op:irs} after rank relaxation can be solved by existing optimization tools such as CVX \cite{Grant2013}. The SCA algorithm of passive beamforming is summarized in Algorithm~\ref{al:sca}.
			\begin{algorithm}[!t]
				\caption{SCA: IRS Phase Shift.}
				\label{al:sca}
				\begin{algorithmic}[1]
					\State \textbf{Input} $\beta_2,\beta_4,\boldsymbol{h}_{\mathrm{D},n},\boldsymbol{V}_{n},\boldsymbol{w}_{\mathrm{I/P},n},\rho,\sigma_n,\bar{R},Q,\epsilon$, $\forall n$
					\State Construct $\boldsymbol{V},\boldsymbol{M},\boldsymbol{M}_n,\boldsymbol{C}_{n},\boldsymbol{C}_{\mathrm{I/P},k}$, $\forall n,k$
					\State \textbf{Initialize} $i \gets 0$, $\boldsymbol{\Phi}^{(0)}$
					\State Get $t_{\mathrm{I/P},k}^{(0)}$, $\forall k$ by \eqref{eq:t_k}
					\Repeat
						\State $i \gets i + 1$
						\State Get $\boldsymbol{\Phi}^{(i)}$ by solving \ref{ob:irs} -- \ref{co:irs_sd}
						\State Update $t_{\mathrm{I/P},k}^{(i)}$, $\forall k$ by \eqref{eq:t_k}
						\State Compute $z^{(i)}$ by \eqref{eq:z_irs}
					\Until $\lvert z^{(i)}-z^{(i-1)} \rvert \le \epsilon$
					\State Set $\boldsymbol{\Phi}^{\star}=\boldsymbol{\Phi}^{(i)}$
					\If{$\mathrm{rank}(\boldsymbol{\Phi}^{\star})=1$}
						\State Get $\bar{\boldsymbol{\phi}}^\star$ by eigen decomposition, $\boldsymbol{\Phi}^{\star}=\bar{\boldsymbol{\phi}}^\star(\bar{\boldsymbol{\phi}}^\star)^H$
					\Else
						\State Get $\boldsymbol{U},\boldsymbol{\Lambda}$ by eigen decomposition, $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Lambda}\boldsymbol{U}^H$
						\State Generate $\boldsymbol{\xi}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, $\forall q$
						\State Construct $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Lambda}^{1/2}\boldsymbol{\xi}_q\right)},\boldsymbol{\Phi}_q=\bar{\boldsymbol{\phi}}_q\bar{\boldsymbol{\phi}}_q^H$, $\forall q$
						\State Set $q^{\star}=\arg\max_q{z(\boldsymbol{\Phi}_q)},\bar{\boldsymbol{\phi}}^\star=\bar{\boldsymbol{\phi}}_{q^{\star}}$
					\EndIf
					\State Set $\theta_l^\star=\arg(\bar{\phi}_l^\star/\bar{\phi}_{L+1}^\star)$, $\forall l$, $\boldsymbol{\phi}^{\star}=[e^{j\theta_1^\star},\dots,e^{j\theta_L^\star}]^H$
					\State \textbf{Output} $\boldsymbol{\phi}^{\star}$
				\end{algorithmic}
			\end{algorithm}

@@

			\begin{algorithm}[!t]
				\caption{Modified: IRS Phase Shift.}
				\label{al:lc}
				\begin{algorithmic}[1]
					\State \textbf{Input} $\beta_2,\beta_4,\boldsymbol{h}_{\mathrm{D},n},\boldsymbol{V}_{n},\boldsymbol{w}_{\mathrm{I/P},n},\rho,\sigma_n,Q,\epsilon$, $\forall n$
					\State Construct $\boldsymbol{V},\boldsymbol{M},\boldsymbol{M}_n,\boldsymbol{C}_{n},\boldsymbol{C}_{\mathrm{I/P},k}$, $\forall n,k$
					\State \textbf{Initialize} $i \gets 0$, $\boldsymbol{\Phi}^{(0)}$
					\If{$\rho=0$}
						\State Get $\boldsymbol{\Phi}^{\star}$ by maximizing \eqref{eq:R_irs} w.r.t. \eqref{co:irs_modulus}, \eqref{co:irs_sd}
					\Else
						\Repeat
							\State $i \gets i + 1$
								\State Set $\bar{R}=0$
								\State Get $\boldsymbol{\Phi}^{(i)}$ by solving \ref{ob:irs} -- \ref{co:irs_sd}
								\State Update $t_{\mathrm{I/P},k}^{(i)}$, $\forall k$ by \eqref{eq:t_k}
								\State Compute $z^{(i)}$ by \eqref{eq:z_irs}
						\Until $\lvert z^{(i)}-z^{(i-1)} \rvert \le \epsilon$
						\State Set $\boldsymbol{\Phi}^{\star}=\boldsymbol{\Phi}^{(i)}$
					\EndIf
					\If{$\mathrm{rank}(\boldsymbol{\Phi}^{\star})=1$}
						\State Get $\bar{\boldsymbol{\phi}}^\star$ by eigen decomposition, $\boldsymbol{\Phi}^{\star}=\bar{\boldsymbol{\phi}}^\star(\bar{\boldsymbol{\phi}}^\star)^H$
					\Else
						\State Get $\boldsymbol{U},\boldsymbol{\Lambda}$ by eigen decomposition, $\boldsymbol{\Phi}^{\star}=\boldsymbol{U}\boldsymbol{\Lambda}\boldsymbol{U}^H$
						\State Generate $\boldsymbol{\xi}_q \sim \mathcal{CN}(\boldsymbol{0},\boldsymbol{I}_{L+1})$, $\forall q$
						\State Construct $\bar{\boldsymbol{\phi}}_q=e^{j\arg\left(\boldsymbol{U}\boldsymbol{\Lambda}^{1/2}\boldsymbol{\xi}_q\right)},\boldsymbol{\Phi}_q=\bar{\boldsymbol{\phi}}_q\bar{\boldsymbol{\phi}}_q^H$, $\forall q$
						\State Set $q^{\star}=\arg\max_q{R(\boldsymbol{\Phi}_q)}$ or $q^{\star}=\arg\max_q{z(\boldsymbol{\Phi}_q)}$, $\bar{\boldsymbol{\phi}}^\star=\bar{\boldsymbol{\phi}}_{q^{\star}}$
					\EndIf
					\State Set $\theta_l^\star=\arg(\bar{\phi}_l^\star/\bar{\phi}_{L+1}^\star)$, $\forall l$, $\boldsymbol{\phi}^{\star}=[e^{j\theta_1^\star},\dots,e^{j\theta_L^\star}]^H$
					\State \textbf{Output} $\boldsymbol{\phi}^{\star}$
				\end{algorithmic}
			\end{algorithm}

			Similar to Algorithm~\ref{al:bcd}, we iteratively update the passive beamforming $\boldsymbol{\phi}$ by Algorithm~\ref{al:lc}, the active beamforming $\boldsymbol{b}_{\mathrm{I/P},n}$, $\forall n$ by equation \eqref{eq:b}, and the waveform amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ by \eqref{eq:wf}, \eqref{eq:smf} for TS or \eqref{eq:s_i}, \eqref{eq:s_p} for PS, until convergence. Similar to Proposition \ref{pr:bcd}, strict convergence and local optimality of the low-complexity BCD algorithm are guaranteed if $\boldsymbol{\Phi}$ is rank-\num{1}.
		\end{subsection}
	\end{section}

@@

		We first provide numerical proof on the strict convergence and local optimality of the proposed BCD algorithms. Too see this, we define the eigen ratio as $\nu=\max_l\lambda_l(\boldsymbol{\Phi})/\sum_l\lambda_l(\boldsymbol{\Phi})$ where $\lambda_l(\boldsymbol{\Phi})$ is the $l$-th eigenvalue of $\boldsymbol{\Phi}$. Table~ shows the minimum value of $\nu$ during both BCD algorithms for all tested channel realizations.

		evaluate the rank of IRS matrix $\boldsymbol{\Phi}^\star$ obtained by Algorithm~\ref{al:sca} and \ref{al:lc}.

		We first evaluate the performance of Algorithm~\ref{al:sca} under SDR. It is demonstrated that $\boldsymbol{\Phi}^{\star}$ is rank-\num{1} for all tested channel realizations with different $M$, $N$ and $L$, such that $\boldsymbol{\phi}^{\star}$ can be directly obtained through eigen decomposition and the Gaussian randomization step is not required. Therefore, we claim Algorithm~\ref{al:sca} converges to stationary points of problem~\eqref{op:irs} without performance loss.
